Paper_notes

### 2017-11-19 Paper_notes.Rmd
Add files via upload
papers Nov20

Lipshutz, R. J., Fodor, S. P., Gingeras T. R., and Lockhart D. J.. "High Density Synthetic Oligonucleotide Arrays." Nature Genetics 21, no. 1 Suppl (January 1999): 20-24. doi:10.1038/4447. https://www.ncbi.nlm.nih.gov/pubmed/9915496

@@

Altman, N., Krzywinski M. “Points of Significance: Sources of Variation.” Nature Methods 12, no. 1 (December 30, 2014): 5–6. doi:10.1038/nmeth.3224

@@

Bolstad BM, Irizarry RA, Astrand M, Speed TP. “A Comparison of Normalization Methods for High Density Oligonucleotide Array Data Based on Variance and Bias.” Bioinformatics (Oxford, England) 19, no. 2 (January 22, 2003): 185–93.

@@

Cleveland, William S, Devlin S. “Locally Weighted Regression: An Approach to Regression Analysis by Local Fitting.” Journal of the American Statistical Association 83, no. 403 (1988): 596–610.

@@

>>>>

- Lipshutz, R. J., Fodor, S. P., Gingeras T. R., and Lockhart D. J.. "High Density Synthetic Oligonucleotide Arrays." Nature Genetics 21, no. 1 Suppl (January 1999): 20-24. doi:10.1038/4447. https://www.ncbi.nlm.nih.gov/pubmed/9915496

@@

- Altman, N., Krzywinski M. “Points of Significance: Sources of Variation.” Nature Methods 12, no. 1 (December 30, 2014): 5–6. doi:10.1038/nmeth.3224

@@

- Bolstad BM, Irizarry RA, Astrand M, Speed TP. “A Comparison of Normalization Methods for High Density Oligonucleotide Array Data Based on Variance and Bias.” Bioinformatics (Oxford, England) 19, no. 2 (January 22, 2003): 185–93.

@@

- Cleveland, William S, Devlin S. “Locally Weighted Regression: An Approach to Regression Analysis by Local Fitting.” Journal of the American Statistical Association 83, no. 403 (1988): 596–610.

@@

### Smyth (2004)

- Smyth GK. "Linear Models and Empirical Bayes Methods for Assessing Differential Expression in Microarray Experiments". Statistical Applications in Genetics and Molecular Biology 3 (2004).

Lonnstedt's method to detect differential expression using a B-statistic (2002) is extended to analyze general microarray experiments. The process builds linear models for the experiment where the dependent variable is the gene expression, for example, log2(R)-log2(G); and the idea is to estimate the model coefficient that distinguishes between groups. The hierarchical model assumes a prior distribution for the coefficients and the variances for each gene. A moderated t-statistic is defined as the classic t-statistic but using the posterior variance instead of the sample variance. Under the null hypothesis, the coefficient is zero, and the moderated t-statistic follows a t-distribution where the degrees of freedom are calculated using the data. The paper also proposes the calculation of the posterior odds to measure the evidence for differential expression (B). The authors show the methodology for the estimation of the hyperparameters. Even when the gene identification using the moderate t-statistic or using B would lead similar results, B requires knowledge of more parameters as compared with moderated t. A simulation study shows that the moderated t has lower false discovery rate as compared with other methods under different scenarios.

### Tusher & Tibshirani & Chu (2001)

- Tusher VG., Tibshirani R., and Chu G. "Significance Analysis of Microarrays Applied to the Ionizing Radiation Response"." Proceedings of the National Academy of Sciences of the United States of America 98, no. 9 (April 24, 2001)

Microarrays allow the measurement of thousands of genes in only one experiment but conventional statistical methods to assess differential expression present high false discovery rate (FDR).  Significance Analysis of Microarrays (SAM) is a proposed approach to identify significant changes in gene expression while accounting for multiple comparisons. First, the proposed statistic, d(i), is a relative difference in the expresión divided by the standard desviación of the repeated measures, adding a positive constant in the denominator to adjust for small variability in gene expression. This statistic is calculated for each gene.

Then, permutations are performed, and the statistic is calculated for each permutation getting the distribution of the statistic and the expected relative difference [de(i)].  The d(i) and de(i) are compared, usually plotted together, and thresholds are specified to identify induced or repressed genes. One important thing about this approach is that the thresholds allow an asymmetric identification.  This approach has shown lower FDR as compared with fold change identification methods.

### Goeman & Solari (2014)

- Goeman JJ, Solari A. "Multiple Hypothesis Testing in Genomics". Statistics in Medicine 33, no. 11 (May 20, 2014).

Multiple testing and correction for it are discussed. Three different perspectives for error control are described. Familywise error (FWER) that controls the probability of having at least one error in a set; the false discovery rate (FDR) that controls the expected proportion of errors among the rejections; and the false discovery proportion (FDP) that estimates the proportion of false rejections among the rejections.

The FWER methods:

- Bonferroni which is simple but conservative especially in cases such as in gene expression where the p-values are correlated.

- Holm that is less conservative than Bonferroni by allowing an increasing alpha as less hypothesis remains to be tested.

- Hochberg and Hommel assume positive dependence through stochastic ordering (PDS), which makes them less conservative as compared with Holm.

- Westfall & Young is a sequential permutation method that rejects according to the quartiles and then recalculated the distribution until no more hypothesis are rejected.

The FDR methods:

- Benjamini & Hochberg

- False discovery rate control under general dependence

- False discovery rate control by resampling

- Use of false discovery rate control

The FDP methods:

- Storey’s approach and significance analysis of microarrays

- Efron’s approach

- False discovery proportion confidence

- Meinshausen’s permutation method
